The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-09-04 Paper.Rmd
modified

\newpage
# Introduction

@@

Basically, we want to address the following questions. 1) Do datasets in ecology and related fields have non-Normal copula srtructure? Do they show tail dependence distinct from that of a Normal copula, and in particular do they show asymmetric tail dependence? 2) If so, what are some possible causes of non-Normal copula structure? And 3) what are the consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications? 
In \nameref{Material&Methods}, we described briefly about each datasets we used to answer these questions and provided a detailed descriptions of implemented methods. In the \nameref{Results} section, we summarized our finidings for each datasets. Next two sections are focused on possible causes and consequences of our analysis. Finally, we concluded in the \nameref{Discussion} section.

@@

For bivariate analysis we used four datasets which are briefly described as follows;
to begin with **Soil C and N data**, to explore the correlation of soil organic C and total soil N across a wide geographic range (conterminous US), ecosystem type, and traceable methodology we used the RaCA (Rapid Carbon Assessment)
database [@Wills2014, **or Wills and Loecke pending???**]. Specifically, the soil organic C and total soil N
stocks (Mg C or N m$^{-2}$ ) to 1m of soil depth from the central pedon of the 5930 sites were included in this
study.

Next, the [**BMR data**](https://github.com/sghosh89/BIVAN/blob/master/Data/BMR/bmr-contrasts.csv) contains the trait contrasts from phylogenetically independent contrasts [@SukumaranH2010] of data reported by [@GenoudIM2017] on the basal metabolic rate (BMR) and mass of mammals species. Specifically, the "SELECT" BMR data in their database were used for those species for which data were available. A phylogeny in NEXUS [@NEXUS] format was extracted from the supplementary information of [@GenoudIM2017]. Version 4.4.0 of the DendroPy library [@SukumaranH2010] was used to calculate phylogenetic independent contrasts using Python 2.7. For each contrast, the path length spanned by the contrast (sum of the branch lengths) and the contrast for the mass, natural logarithm of the BMR and natural logarithm of the mass were written as a row in the bmr-contrasts.csv file.

@@

For multivariate analysis, we applied four datasets; first two data sets [for aphid-count and aphid-firstflight date] were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014]. These two annual timesries (1976-2010) measurements are from 11 locations acroos UK for 20 aphid species. We used the same data format as used by @Sheppard2016 for their analysis on the impact of climate on aphids' spatial synchrony. In our case, we chose species=10 [Name : "Green spruce aphid"] for aphid-abundance data and species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data and considered only those locations which had atleast 30 years of data (see Figure \ref{SM-fig_aphid_map}). Figure \ref{fig_aphid_raw_cop} represents raw and copula data plot for aphid count ("Green spruce aphid") and aphid-firstflight date ("Leaf-curling plum aphid") taken from two locations.

@@

Question 1 from the Introduction was answered via statistical analysis of the
data described above, using both model selection and nonparametric approaches.

@@

A model-selection approach based on likelihood and the Akaike Information Criterion (AIC) 
and Bayesian Information Criterion (BIC) was used to
help answer question 1 from the Introduction.

@@

The first part or question 1 from the Introduction was answered for a bivariate dataset 
$(x_i,y_i)$ for $i=1,...,n$ by comparing the AIC weights of the non-Normal copula
families with that of the Normal family. The same was also done using BIC.
The second part of question 1 was answered by comparing 
the model-averaged lower- and upper-tail dependence
values to the lower- and upper-tail dependence values of a Normal copula, both of which 
are 0. The third part of question 1
was answered by comparing the difference of the model-averaged lower- and upper-tail 
dependence values to the analogous difference for a Normal copula, again 0.

@@

In addition to the model selection approach described above, nonparametric 
approaches were also used to answer question 1 from the introduction.
Using two distinct methods, when results are consistent, provides greater confidence 
that results are robust.

@@

The comparison of one of the
statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
empirical data, to the distribution
of its values computed on many surrogate datasets 
provides a test of whether dependence in the part of the distribution specified 
by $l_b$ and $u_b$ is different from
what would be expected from a null hypothesis of normal copula structure but 
unchanged marginals and unchanged 
Kendall (or Spearman)
correlation. Thus the comparison helps answer the first two parts of question 1 
from the Introduction for bivariate datasets. 
Significant deviations correspond to deviations from normal 
copula structure. In particular, deviations 
using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence 
different from that of a normal copula;
likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different 
from that of a normal copula.

The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, 
as computed on the 
empirical data, to the distribution
of its values on many surrogate datasets 
provides a test of asymmetry of tail dependence. Thus this comparison helps answer the
third part of question 1 from the Introduction for bivariate datasets.

For multivariate datasets, a spatial resampling scheme was used instead of the surrogate 
approach described above.
The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 
as part of a method for 
performing model selection among matrix regression models. For each pair of 
locations $i \neq j$, the statistic
of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
values $s_{ij}$ were then arranged in
a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the 
average strength of 
dependence, across all pairs of distinct locations, for data in the portion of the distributions 
given by $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, the mean characterized average
asymmetry of tail dependence across all pairs of locations. 
Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a 
distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
revealed by these confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
Normal null hypothesis, and answering question 1 from the Introduction.
Bootstrapping was performed by resampling the locations of measurement, with replacement. 
This corresponds to 
resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
[@bjornstadfalck2001 ; @walteretal2017].

@@

## question 1

- Do datasets in ecology and related fields have non-Normal copula structure?
- Do they show tail dependence distinct from that of a Normal copula?
- In particular do they show asymmetric tail dependence?

To answer Q1, we choose one representative datasets for each category : with a bivariate (soil C and N data) and
a multivariate (green spruce aphid-abundance data) dataset, we summarized our findings as follows.

### soil C and N data \label{Ex_SoilCN}

@@

As described in the \ref{SM-SoilCN_data}, raw data plot for soil organic C and soil N stocks to 1m of soil depth from the central
pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig. \ref{fig_biv_raw_cop}(A) and (D),respectively.
We carried out model selection analysis and non-parametric statistics to measure tail-dependence (if any) on this copula plot.

@@

The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes` (to within the precision available 
from `BiCopIndTest` in the `VineCopula` package) that showed a significant correlation (Kendall's coefficient = `r round(BivMS_res_Loecke$TauVal,2)`). We then fitted several copulas (all $16$ families as mentioned in \nameref{Model_selection})
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_soilCNfit}). 
Based on AIC weights, `r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the best fit. The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of question 1 for these data.

The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`. These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of question 1. To be precise, this difference being negative, indicated stronger upper-tail dependence than lower.

@@

Results indicate that tail dependence of data is stronger in both the lower and upper tails than would be expected under a 
null hypothesis of a normal copula, answering the second part of question 1. The dependence between C and N is also 
weaker in the middle of the distributions
than would be expected based on a normal-copula null model.

We also tested statistically for asymmetry of tail dependence, which seems visible in Fig. \ref{SM-fig_raw_cop_soilCN}(B). 
Results (Table \ref{SM-tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of question 1.

We summarized our results columnwise in Table \ref{tab_bivar_summary} for all bivariate datasets plotted in Fig. \ref{fig_biv_raw_cop}.

The first $9$ rows of Table \ref{tab_bivar_summary} summarized model selection results [namely, (1) p-value from `BiCopIndTest`, (2) best copula based on minimum AIC, (3) AIC-weight of that best copula, (4) AIC-weight of Normal copula, (5) p-value from `BiCopGofTest` for Cramer-von Mises method, (6) p-value from `BiCopGofTest` for Kolmogorov-Smirnov method, (7) AIC-weights model averaged lower tail dependence, (8) AIC-weights model averaged upper tail dependence, (9)  AIC-weights model averaged (lower-upper) tail dependence].
Last $3$ rows ($10^{th}$ to $12^{th}$) of Table \ref{tab_bivar_summary} summarized non-parametric statistics results for $P$ statistics calculated on two extreme bins ($\Ps_l$ for first bin, $\Ps_u$ for last bin) for each copula data. 
We computed $\Ps_l$, $\Ps_u$, $\Ps_l - \Ps_u$ for the actual copula data and for Kendall-correlation 
preserving $1000$ Normal surrogates and calculated the rank for actual data as the position if all these $1001$ values were 
arranged in ascending order. So, whenever those ranks are $>975$ or $<25$, that indicate a significantly ($95\%$ CI) different 
tail-dependence measures based on Normal-null hypothesis.

For all $4$ bivariate-datasets indicated by each column name (SoilCN, BMR, Anolis, Ceder-creek) 
in Table \ref{tab_bivar_summary}, Normal copula were always a poor fit compared to 
highest AIC-weight based asymmetric copulas - which answers the first part of Q1. 
Among four datasets presented there, first three have significant upper tail dependence whereas the last one has significant lower tail dependence compared to Normal. It was indicated by negative AIC-weights model averaged 
(lower-upper) tail dependence and very low ranks of actual copula 
data for $\Ps_l - \Ps_u$ statistics for first three datasets and positive AIC-weights model averaged 
(lower-upper) tail dependence and very high ranks of actual copula 
data for $\Ps_l - \Ps_u$ statistics for the last one, which answers the second and third part of Q1
(For detailed results of each dataset see Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} and 
Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}).

@@

### Abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}

@@

As described in \nameref{Model_selection}, for multivariable dataset like green-spruce aphid abundance data, model selection among copulas was done separately for each pair of the 10 sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns corresponding to locations. Table \ref{SM-tab_count_bestcop} shows that, for the large majority of location pairs, some non-Normal copula was a better fit, according to maximum AIC-weights, than was the Normal copula, answering the first part of question 1 for these data. Comparing the 
AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_count_AICw}) to the AIC weight of the Normal copula family for the same pair (Table \ref{SM-tab_count_AICw_normal}) confirms that the degree of support for Normal copulas was usually not
comparable to the degree of support for other families.
Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_count_p_CvM}) in every case 
failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_count_p_KS}), demonstrating that the collection of copula families we used was sufficiently broad that the copulas of all pairs of locations could be characterized by at least one of them. In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.

The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are in Tables \ref{SM-tab_count_LTdep_AICw} and \ref{SM-tab_count_UTdep_AICw} respectively. These values were exclusively positive, thereby differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data. Apparently these aphid data have greater tail dependence (both lower and upper) than would be expected from a normal-copula null hypothesis.

The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{SM-tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question 1. Fig. \ref{fig_aphid_raw_cop}(A) and (C) represnt exemplary plots for aphid abundance (green spruce aphid) raw data and its copula form at $(4,11)$ location pair which shows stronger lower tail dependence.

@@

Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
resampling scheme, as described in \nameref{NPA} (Table \ref{SM-tab_count_resamp}). Confidence intervals of
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
exluded $0$ and were positive, indicating asymmetry of tail dependence (stronger lower tail dependence than upper), a feature that data with Normal copula structure do not have, providing further evidence answering the third part of question 1.

@@

<!--Question 2 analyses-->
## Some possible causes of non-Normal copula structure

@@

<!--Question 3 analyses-->
## Consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications?

>>>>

\newpage
# Introduction \label{Intro}

@@

Basically, we want to address the following questions. \textbf{(Q1)} Do datasets in ecology and related fields have non-Normal copula srtructure? Do they show tail dependence distinct from that of a Normal copula, and in particular do they show asymmetric tail dependence? \textbf{(Q2)} If so, what are some possible causes of non-Normal copula structure? And \textbf{(Q3)} what are the consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications? 
In \nameref{Material&Methods}, we described briefly about each bivariate and multivariate datasets we used to answer these questions and provided a detailed descriptions of implemented methods. In the \nameref{Results} section, we summarized our finidings for each datasets. Next two sections and are focused on possible causes and consequences of non-Normal copula structure in ecology and related field. Finally, we concluded in the \nameref{Discussion} section.

@@

For bivariate analysis we used four datasets which are briefly described as follows;
First, to explore the dependence of **soil organic C** and total **soil N** across a wide geographic range (conterminous US), ecosystem type, and traceable methodology we used the RaCA (Rapid Carbon Assessment)
database [@Wills2014, **or Wills and Loecke pending???**]. Specifically, the soil organic C and total soil N
stocks (Mg C or N m$^{-2}$ ) to 1m of soil depth from the central pedon of the 5930 sites were included in this
study.

Next, the [**BMR data**](https://github.com/sghosh89/BIVAN/blob/master/Data/BMR/bmr-contrasts.csv) contains the trait contrasts from phylogenetically independent contrasts [@SukumaranH2010] of data reported by [@GenoudIM2017] on the basal metabolic rate (BMR) and mass of mammals species. Specifically, the "SELECT" BMR data in their database were used for those species for which data were available. A phylogeny in NEXUS [@NEXUS] format was extracted from the supplementary information of [@GenoudIM2017]. Version 4.4.0 of the DendroPy library [@SukumaranH2010] was used to calculate phylogenetic independent contrasts using Python 2.7. For each contrast, the path length spanned by the contrast (sum of the branch lengths) and the contrast for the mass, natural logarithm of the BMR and natural logarithm of the mass were written as a row in the [bmr-contrasts.csv](https://github.com/sghosh89/BIVAN/blob/master/Data/BMR/bmr-contrasts.csv) file.

@@

Similarly, we applied multivariate analysis on four datasets; first two data sets [for aphid-count and aphid-firstflight date] were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014]. These two annual timesries (1976-2010) measurements are from 11 locations acroos UK for 20 aphid species. We used the same data format as used by @Sheppard2016 for their analysis on the impact of climate on aphids' spatial synchrony. In our case, we chose species=10 [Name : "Green spruce aphid"] for aphid-abundance data and species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data and considered only those locations which had atleast 30 years of data (see Figure \ref{SM-fig_aphid_map}). Figure \ref{fig_aphid_raw_cop} represents raw and copula data plot for aphid count ("Green spruce aphid") and aphid-firstflight date ("Leaf-curling plum aphid") taken from two locations.

@@

Question \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro} was answered via statistical analysis of the
data described above, using both model selection and nonparametric approaches.

@@

A model-selection approach based on likelihood and the Akaike Information Criterion (AIC) 
and Bayesian Information Criterion (BIC) was used to
help answer \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro}.

@@

The first part of \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro} was answered for a bivariate dataset 
$(x_i,y_i)$ for $i=1,...,n$ by comparing the AIC weights of the non-Normal copula
families with that of the Normal family. The same was also done using BIC.
The second part of \hyperref[Intro]{\textbf{(Q1)}} was answered by comparing 
the model-averaged lower- and upper-tail dependence
values to the lower- and upper-tail dependence values of a Normal copula, both of which 
are 0. The third part of \hyperref[Intro]{\textbf{(Q1)}}
was answered by comparing the difference of the model-averaged lower- and upper-tail 
dependence values to the analogous difference for a Normal copula, again 0.

@@

In addition to the model selection approach described above, nonparametric 
approaches were also used to answer \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro}.
Using two distinct methods, when results are consistent, provides greater confidence 
that results are robust.

@@

The comparison of one of the
statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
empirical data, to the distribution
of its values computed on many surrogate datasets 
provides a test of whether dependence in the part of the distribution specified 
by $l_b$ and $u_b$ is different from
what would be expected from a null hypothesis of normal copula structure but 
unchanged marginals and unchanged 
Kendall (or Spearman)
correlation. Thus the comparison helps answer the first two parts of question \hyperref[Intro]{\textbf{(Q1)}} 
from the \nameref{Intro} for bivariate datasets. 
Significant deviations correspond to deviations from normal 
copula structure. In particular, deviations 
using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence 
different from that of a normal copula;
likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different 
from that of a normal copula.

The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, 
as computed on the 
empirical data, to the distribution
of its values on many surrogate datasets 
provides a test of asymmetry of tail dependence. Thus this comparison helps answer the
third part of question \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro} for bivariate datasets.

For multivariate datasets, a spatial resampling scheme was used instead of the surrogate 
approach described above.
The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 
as part of a method for 
performing model selection among matrix regression models. For each pair of 
locations $i \neq j$, the statistic
of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
values $s_{ij}$ were then arranged in
a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the 
average strength of 
dependence, across all pairs of distinct locations, for data in the portion of the distributions 
given by $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, the mean characterized average
asymmetry of tail dependence across all pairs of locations. 
Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a 
distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
revealed by these confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
Normal null hypothesis, and answering question \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro}.
Bootstrapping was performed by resampling the locations of measurement, with replacement. 
This corresponds to 
resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
[@bjornstadfalck2001 ; @walteretal2017].

@@

<!--## question 1

- Do datasets in ecology and related fields have non-Normal copula structure?
- Do they show tail dependence distinct from that of a Normal copula?
- In particular do they show asymmetric tail dependence?-->

To answer \hyperref[Intro]{\textbf{(Q1)}}, we choose one representative datasets for each category : with a bivariate (soil C and N data) and a multivariate (green spruce aphid-abundance data) dataset, we summarized our findings as follows.

### Example for bivariate case : soil C and N data \label{Ex_SoilCN}

@@

As described in the \nameref{Data}, raw data plot for soil organic C and soil N stocks to 1m of soil depth from the central pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig. \ref{fig_biv_raw_cop}(A) and (E),respectively.
We carried out model selection analysis and non-parametric statistics to measure tail-dependence (if any) on this copula plot.

@@

The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes` (to within the precision available 
from `BiCopIndTest` in the `VineCopula` package) that showed a significant correlation (Kendall's coefficient = `r round(BivMS_res_Loecke$TauVal,2)`). We then fitted several copulas (all $16$ families as mentioned in \nameref{Model_selection})
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_soilCNfit}). 
Based on AIC weights, `r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the best fit. The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of \hyperref[Intro]{\textbf{(Q1)}} for these data.

The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`. These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of \hyperref[Intro]{\textbf{(Q1)}} for these data. Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of \hyperref[Intro]{\textbf{(Q1)}}. To be precise, this difference being negative, indicated stronger upper-tail dependence than lower.

@@

Results indicate that tail dependence of data is stronger in both the lower and upper tails than would be expected under a 
null hypothesis of a normal copula, answering the second part of \hyperref[Intro]{\textbf{(Q1)}}. The dependence between C and N is also 
weaker in the middle of the distributions
than would be expected based on a normal-copula null model.

We also tested statistically for asymmetry of tail dependence, which seems visible in Fig. \ref{fig_biv_raw_cop}(E). 
Results (Table \ref{SM-tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of \hyperref[Intro]{\textbf{(Q1)}}.

We summarized our results columnwise in Table \ref{tab_bivar_summary} for all bivariate datasets that were plotted in Fig. \ref{fig_biv_raw_cop}.

The first $9$ rows of Table \ref{tab_bivar_summary} summarized model selection results [namely, (1) p-value from `BiCopIndTest`, (2) best copula based on minimum AIC, (3) AIC-weight of that best copula, (4) AIC-weight of Normal copula, (5) p-value from `BiCopGofTest` for Cramer-von Mises method, (6) p-value from `BiCopGofTest` for Kolmogorov-Smirnov method, (7) AIC-weights model averaged lower tail dependence, (8) AIC-weights model averaged upper tail dependence, (9)  AIC-weights model averaged (lower-upper) tail dependence].
Last $3$ rows ($10^{th}$ to $12^{th}$) of Table \ref{tab_bivar_summary} summarized non-parametric statistics results for $P$ statistics calculated on two extreme bins ($\Ps_l$ for first bin, $\Ps_u$ for last bin) for each copula data. 
We computed $\Ps_l$, $\Ps_u$, $\Ps_l - \Ps_u$ for the actual copula data and for Kendall-correlation 
preserving $1000$ Normal surrogates and calculated the rank for actual data as the position if all these $1001$ values were 
arranged in ascending order. So, whenever those ranks were $>975$ or $<25$, that indicated a significantly ($95\%$ CI) different 
tail-dependence measures based on Normal-null hypothesis.

For all $4$ bivariate-datasets indicated by each column name (SoilCN, BMR, Anolis, Ceder-creek) 
in Table \ref{tab_bivar_summary}, Normal copula were always a poor fit compared to 
highest AIC-weight based asymmetric copulas - which answers the first part of \hyperref[Intro]{\textbf{(Q1)}}. 
Among four datasets presented there, first three have significant upper tail dependence whereas the last one has significant lower tail dependence compared to Normal. It was indicated by negative AIC-weights model averaged 
(lower-upper) tail dependence and very low ranks of actual copula 
data for $\Ps_l - \Ps_u$ statistics for first three datasets and positive AIC-weights model averaged 
(lower-upper) tail dependence and very high ranks of actual copula 
data for $\Ps_l - \Ps_u$ statistics for the last one, which answers the second and third part of \hyperref[Intro]{\textbf{(Q1)}}
(For detailed results of each dataset see Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} and 
Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}).

@@

### Example for multivariate case : abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}

@@

As described in \nameref{Model_selection}, for multivariable dataset like green-spruce aphid abundance data, model selection among copulas was done separately for each pair of the 10 sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns corresponding to locations. Table \ref{SM-tab_count_bestcop} shows that, for the large majority of location pairs, some non-Normal copula was a better fit, according to maximum AIC-weights, than was the Normal copula, answering the first part of question \hyperref[Intro]{\textbf{(Q1)}} for these data. Comparing the 
AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_count_AICw}) to the AIC weight of the Normal copula family for the same pair (Table \ref{SM-tab_count_AICw_normal}) confirms that the degree of support for Normal copulas was usually not
comparable to the degree of support for other families.
Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_count_p_CvM}) in every case 
failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_count_p_KS}), demonstrating that the collection of copula families we used was sufficiently broad that the copulas of all pairs of locations could be characterized by at least one of them. In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.

The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are in Tables \ref{SM-tab_count_LTdep_AICw} and \ref{SM-tab_count_UTdep_AICw} respectively. These values were exclusively positive, thereby differing from what a normal copula would give (i.e., 0), and answering the second part of question \hyperref[Intro]{\textbf{(Q1)}} for these data. Apparently these aphid data have greater tail dependence (both lower and upper) than would be expected from a normal-copula null hypothesis.

The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{SM-tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question \hyperref[Intro]{\textbf{(Q1)}}. Fig. \ref{fig_aphid_raw_cop}(A) and (C) represnt exemplary plots for aphid abundance (green spruce aphid) raw data and its copula form at $(4,11)$ location pair which shows stronger lower tail dependence.

@@

Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
resampling scheme, as described in \nameref{NPA} (Table \ref{SM-tab_count_resamp}). Confidence intervals of
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
exluded $0$ and were positive, indicating asymmetry of tail dependence (stronger lower tail dependence than upper), a feature that data with Normal copula structure do not have, providing further evidence answering the third part of question \hyperref[Intro]{\textbf{(Q1)}}.

@@

<!--Question 2 analyses-->
## Some possible causes of non-Normal copula structure \label{Causes}

@@

<!--Question 3 analyses-->
## Consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications? \label{Consequences}
